ppseq: An R Package for Sequential Predictive Probability Monitoring

### 2022-11-15 inst/ppseq-package-paper/zabor-hobbs-kane.Rmd
final paper version

We wish to re-design the phase II trial of atezolizumab using a two-arm randomized design with sequential predictive probability monitoring. We willuse the `calibrate_thresholds()` function to obtain the operating characteristics of designs based on on each combination of posterior and predictive thresholds. We focus here on the main biomarker subgroup of interest, the IC2/3 subgroup. We design the study with a null response rate of 0.1 in both arms, and an alternative response rate of 0.25 in the atezolizumab arm. So we set `p_null = c(0.1, 0.1)` and `p_alt = c(0.1, 0.25)`. We plan the study with 100 participants, assuming that the total sample size available is similar to the 310 used in the actual single-arm phase II trial, and that a third of that patient population fall into our desired biomarker subgroup. So we set `N = c(50, 50)` to indicate equal allocation of the 100 patients to each of the two arms. We will check for futility after every 10 patients are enrolled on each arm, so we set `n = cbind(seq(10, 50, 10), seq(10, 50, 10))`, a matrix showing the interim analysis sample sizes in each of the two arms. To design the study, we need to calibrate the design over a range of posterior probability thresholds and predictive probability thresholds. In this example we will consider posterior thresholds of 0.9, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96, 0.97, 0.98, and 0.99, and predictive thresholds of 0.05, 0.1, 0.15, and 0.2. So we set `pp_threshold = seq(0.9, 0.99, 0.01)` and `ppp_threshold = seq(0.05, 0.2, 0.05)`. We selected these posterior thresholds because we want to have a posterior probability of at least 0.9 that our response rate exceeds the null. And we selected these predictive thresholds because we would  not want to stop the trial early for futility if there was more than a 20\% chance of success when full enrollment was reached. We use the defaults for all remaining arguments, including the defaults of `S=5000` and `nsim=1000`, which we believe are sufficient to obtain stable results. Because of the inherent computational intensity in these calculations, this function relies on the \CRANpkg{future} [@Bengtsson2021] and \CRANpkg{furrr} [@Vaughan2021] packages to parallelize computations. The user will be responsible for setting up a call to `future::plan()` that is appropriate to their operating environment and simulation setting. Because the code takes some time to run, the results of the below example code are available as a dataset called `two_sample_cal_tbl` included in the \CRANpkg{ppseq} package.

>>>>

We wish to re-design the phase II trial of atezolizumab using a two-arm randomized design with sequential predictive probability monitoring. We will use the `calibrate_thresholds()` function to obtain the operating characteristics of designs based on on each combination of posterior and predictive thresholds. We focus here on the main biomarker subgroup of interest, the IC2/3 subgroup. We design the study with a null response rate of 0.1 in both arms, and an alternative response rate of 0.25 in the atezolizumab arm. So we set `p_null = c(0.1, 0.1)` and `p_alt = c(0.1, 0.25)`. We plan the study with 100 participants, assuming that the total sample size available is similar to the 310 used in the actual single-arm phase II trial, and that a third of that patient population fall into our desired biomarker subgroup. So we set `N = c(50, 50)` to indicate equal allocation of the 100 patients to each of the two arms. We will check for futility after every 10 patients are enrolled on each arm, so we set `n = cbind(seq(10, 50, 10), seq(10, 50, 10))`, a matrix showing the interim analysis sample sizes in each of the two arms. To design the study, we need to calibrate the design over a range of posterior probability thresholds and predictive probability thresholds. In this example we will consider posterior thresholds of 0.9, 0.91, 0.92, 0.93, 0.94, 0.95, 0.96, 0.97, 0.98, and 0.99, and predictive thresholds of 0.05, 0.1, 0.15, and 0.2. So we set `pp_threshold = seq(0.9, 0.99, 0.01)` and `ppp_threshold = seq(0.05, 0.2, 0.05)`. We selected these posterior thresholds because we want to have a posterior probability of at least 0.9 that our response rate exceeds the null. And we selected these predictive thresholds because we would  not want to stop the trial early for futility if there was more than a 20\% chance of success when full enrollment was reached. We use the defaults for all remaining arguments, including the defaults of `S=5000` and `nsim=1000`, which we believe are sufficient to obtain stable results. Because of the inherent computational intensity in these calculations, this function relies on the \CRANpkg{future} [@Bengtsson2021] and \CRANpkg{furrr} [@Vaughan2021] packages to parallelize computations. The user will be responsible for setting up a call to `future::plan()` that is appropriate to their operating environment and simulation setting. Because the code takes some time to run, the results of the below example code are available as a dataset called `two_sample_cal_tbl` included in the \CRANpkg{ppseq} package.
